---
title: "Lab 5"
author: "Devon Rossi"
date: "10/26/2021"
output: 
  html_document: 
    theme: cerulean
    number_sections: yes
    code_folding: hide
    toc: yes
---
#Setup
```{r setup, include= TRUE, warning = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


#Confidence Intervals

In this section, we'll learn to use the 't.test()' function as a way to return the confidence interval using t-distribution.

## Example 1

Create a mock sample of observation values, and find the confidence interval. Then we'll learn how to report that.

```{r}
#QUESTION: do I think the sample I've collected is representative of pop I'm studying?
# How are these values distributed & do these values make sense?
# Normal underlying distribution -> time to find CI 

pika_mass <- c(82, 94, 110, 70, 86, 99, 102, 320, 5000)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)

pika_ci

#Run in console (pika_ci)
# Look in Env -> List of values
# t-value: converted mean value from how far away it is from null hypothesis of ZERO on a t-distribution
```

### Risky & Tedious way :o
~ due to human error, doesn't update when we change values ~

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g 
(mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g.

### A high initial investment, but safer and better way :)

In-line code referecing to stored objects!

Use a single backtick on either side of a lowercase r to create an in-line bit of code.

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round (pika_ci$conf.int[1], 1)`, `r round(pika_ci$conf.int[2], 1)`] g.

### Indexing
- To pull a value from pika_mass, for example, go to Console
- To find 94 g, use pika_mass[2] since it is the second value

# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: Means difference = 0
- HA: Means difference in NOT = 0

This is a two-sided t-test because no directionality is applied.

## Example 1

```{r}
# we think a t-test is a responsible way to answer the question: do we think these samples are drawn from a sample w/ the same mean?

oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)

```


Is there a significant difference in mean heights for oak and pine trees?

```{r}
trees_t <- t.test(oak_trees, pine_trees)

```

The p-value of `r trees_t$p.value` means that ...
p-value: 0.01843 

`r (trees_t$p.value) * 100`% chance of finding sample means *at least as different as those I found* by random chance if they were drawn from populations with the same mean height.

On a t-distrubtion, how far away from null hypothesis - probability associated w that happening is not likely (only 1.8% chance) ... low p-value indicates that it is UNLIKELY that trees are taken from populations with the same mean height 

integrating under t-distribution beyond value of t-statistic in either direction

2/100 times -> we would expect to find sample means that are AT LEAST AS DIFF as sample means found - possible, but so unlikely to happen by random chance -> more likely drawn from pops w different means to begin with

probably started off in populations w/ different means to start with so we REJECT the null hypothesis (cutoff is 5%, or p-value of 0.05)


## Example 2

Using the `mpg` dataset to compare city gas mileage for SUVs and compact cars.
